---
title: "lab-03"
author: '[Matthew Coleman](https://matthewbcoleman.github.io/)'
subtitle: 'Lab 03: Distances and the Border Zone'
output: 
  html_document:
    theme: journal
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(dplyr.summarise.inform=F) 
```

```{r, include = FALSE}
# SPDS
library(tidyverse)
library(sf)
library(units)
library(scales)

# Data
library(USAboundaries)
library(rnaturalearth)

# Visualization
library(gghighlight)
library(ggrepel)
library(knitr)
library(ggthemes)
library(kableExtra)
```

# Data

## Projection

```{r}
eqdc = '+proj=eqdc +lat_0=40 +lon_0=-96 +lat_1=20 +lat_2=60 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs'
```

The PRØJ string `r eqdc` represents:

* Projection name is eqdc (equidistant)
* Latitude of origin: 40
* Longitude of origin: -96
* Latitude of first standard parallel: 20
* Latitude of second standard parallel: 60
* False Easting: 0
* False Northing: 0
* Datum: NAD 83
* Units: Meters


```{r}
## USA State Boundaries

nonstate <- c('Puerto Rico', 'Hawaii', 'District of Columbia', 'Alaska')

usstates_raw <- USAboundaries::us_states(resolution = 'low') %>% st_transform(eqdc)

conus <- usstates_raw %>% 
  filter(!(name %in% nonstate))

conus <- st_transform(conus, eqdc)
```



```{r}
## Mexico, Canada, and US Country Borders

country_borders <- countries110 %>% st_as_sf() %>% 
  filter(admin %in% c('Canada', 'Mexico', 'United States of America')) %>%
  st_transform(eqdc)
```



```{r, message = F}
## Cities

cities <- read_csv('../data/uscities.csv') %>%
  st_as_sf(coords = c('lng', 'lat'), crs = 4326) %>%
  filter(!(state_name %in% nonstate)) %>% 
  st_transform(eqdc)
```

# United States Cities Distances

## Distance to USA Border

```{r}
us_boundary <- st_union(usstates_raw) %>%
  st_cast('MULTILINESTRING')
```

```{r}
city_boundary <- cities %>%
    mutate(dist_to_border = st_distance(cities, us_boundary),
           dist_to_border = set_units(dist_to_border, 'km'),
           dist_to_border = drop_units(dist_to_border))

top_5_border <- city_boundary %>% 
  slice_max(dist_to_border, n = 5) %>% 
  select(city, state_name, dist_to_border) %>%
  st_set_geometry(NULL)
```

```{r}
kable(top_5_border, caption = 'Top 5 Cities by Distance to US Border', 
      col.names = c('City','State', 'Distance (km)'), 
      align = 'l',
      format.args = list(big.mark = ',')) %>% 
  kable_styling()
```

## Distance to States

```{r}
state_boundary <- st_combine(usstates_raw) %>% st_cast('MULTILINESTRING')

state_city_boundary <- cities %>%
    mutate(dist_to_border = {st_distance(cities, state_boundary) %>% set_units('km') %>% 
        drop_units()})

state_top_5_border <- state_city_boundary %>% 
  slice_max(dist_to_border, n = 5) %>% 
  select(city, state_name, dist_to_border) %>%
  st_set_geometry(NULL)
```

```{r}
kable(state_top_5_border, caption = 'Top 5 Cities by Distance to State Border', 
      col.names = c('City','State', 'Distance (km)'), 
      align = 'l',
      format.args = list(big.mark = ',')) %>% 
  kable_styling()
```

## Distance to Mexico Border

```{r}
mexico_border <- country_borders %>% filter(admin== 'Mexico') %>% st_union() %>%  st_cast('MULTILINESTRING')

city_mx_boundary <- cities %>%
    mutate(dist_to_mx = {st_distance(cities, mexico_border) %>% set_units('km') %>% 
        drop_units()})

mx_top_5_border <- city_mx_boundary %>% 
  slice_max(dist_to_mx, n = 5) %>% 
  select(city, state_name, dist_to_mx) %>%
  st_set_geometry(NULL)
```

```{r}
kable(mx_top_5_border, caption = 'Top 5 Cities by Distance to Mexican Border', 
      col.names = c('City','State', 'Distance (km)'), 
      align = 'l',
      format.args = list(big.mark = ',')) %>% 
  kable_styling()
```

## Distance to Canada

```{r}
canada_border <- country_borders %>% filter(admin== 'Canada') %>% st_union() %>%  st_cast('MULTILINESTRING')

city_ca_boundary <- cities %>%
    mutate(dist_to_ca = {st_distance(cities, canada_border) %>% set_units('km') %>% 
        drop_units()})

ca_top_5_border <- city_ca_boundary %>% 
  slice_max(dist_to_ca, n = 5) %>% 
  select(city, state_name, dist_to_ca) %>%
  st_set_geometry(NULL)
```

```{r}
kable(ca_top_5_border, caption = 'Top 5 Cities by Distance to Canadian Border', 
      col.names = c('City','State', 'Distance (km)'), 
      align = 'l',
      format.args = list(big.mark = ',')) %>% 
  kable_styling()
```

# Mapping Cities

## Outlines of US, Canada, Mexico, and the US's 10 Largest Cities by Population 

```{r}
big_cities <- cities %>% slice_max(population, n = 10)

top_10_pop <- ggplot() + 
  geom_sf(data = country_borders) +
  geom_sf(data = conus, color = 'red', lty = 2, size = .5) +
  geom_sf(data = big_cities, size = 1) +
  scale_color_gradient(low = 'grey', high = 'red') +
  theme_map() +
  geom_label_repel(data = big_cities,
                   aes(label = city, geometry = geometry),
                   stat = 'sf_coordinates',
                   size = 2.75) +
  labs('Title Map of Canada, Mexico, US, and top 10 US States by Population')
  
top_10_pop
```

## City Distance from US Border

<center>

```{r}
top_us_border <- city_boundary %>% 
  slice_max(dist_to_border, n = 5) 

us_border_plot  <- ggplot() + 
  geom_sf(data = conus) +
  geom_sf(data = city_boundary, aes(col = dist_to_border), size = .05) +
  geom_sf(data = top_us_border)+
  scale_color_gradient(low = 'grey', high = 'red') +
  theme_map() +
  geom_label_repel(data = top_us_border,
                   aes(label = city, geometry = geometry),
                   stat = 'sf_coordinates',
                   size = 2.75) +
  labs(col = 'Distance to Border (km)', 
       title = 'Map of US Cities and Distance to US Border',
       subtitle = 'Top 5 Cities by Distance to US Border Labeled')
  
  

us_border_plot
```
</center>

## Distance to Nearest Border

<center>

```{r}

state_top_5_geom <- state_city_boundary %>% 
  slice_max(dist_to_border, n = 5) 


state_border_plot  <- ggplot() + 
  geom_sf(data = conus) +
  geom_sf(data = state_city_boundary, aes(col = dist_to_border), size = .05) +
  geom_sf(data = state_top_5_geom )+
  scale_color_gradient(low = 'grey', high = 'red') +
  theme_map() +
  geom_label_repel(data = state_top_5_geom ,
                   aes(label = city, geometry = geometry),
                   stat = 'sf_coordinates',
                   size = 2.75) +
  labs(col = 'Distance to Border (km)',
       title = 'Map of US Cities and Distance to State Border',
       subtitle = 'Top 5 Cities by Distance to State Border Labeled')


state_border_plot
```
</center>


## Equidistant Boundary to Mexico and Canada

<center>

```{r}
equi_boundary <- cities %>%
    mutate(dist_to_ca = {st_distance(cities, canada_border) %>% set_units('km') %>% drop_units()},
           dist_to_mx = {st_distance(cities, mexico_border) %>% set_units('km') %>% drop_units()},
           equidistance = {abs(dist_to_ca-dist_to_mx)}
        ) %>% 
  filter(equidistance <= 100)

equi_pop <- equi_boundary %>% 
  slice_max(population, n = 5) 


mx_ca_border_plot  <- ggplot() + 
  geom_sf(data = conus) +
  geom_sf(data = equi_boundary, size = .05, color = 'darkred', alpha = .4) +
  geom_sf(data = equi_pop)+
  scale_color_gradient(low = 'grey', high = 'red') +
  theme_map() +
  geom_label_repel(data = equi_pop ,
                   aes(label = city, geometry = geometry),
                   stat = 'sf_coordinates',
                   size = 2.75) +
  labs(col = 'Distance to Border (km)', 
       title = 'Map of US Cities Equidistant to Canada and Mexico Border ± 100 km',
       subtitle = 'Top 5 Cities by Population Labeled')

mx_ca_border_plot

```
</center>

# 100 Mile (160 km) Border Zone

## Cities Within the Border Zone

```{r}
us_border_100 <- city_boundary %>% filter(dist_to_border <= 160)

number_100_mile <- nrow(us_border_100)

population_100_mile <- sum(us_border_100$population)

proportion_100_mile <- population_100_mile/sum(city_boundary$population)

descriptions <- c('Number of cities within the 100-mile zone', 'Population within the 100 mile zone', 'Proportion of population within the 100-mile zone')
values <- c(number_100_mile, population_100_mile, label_percent()(proportion_100_mile))

table <- data.frame(descriptions, values)

kable(table, caption = '100-Mile Zone Facts', 
      col.names = c('Metric','Value'), 
      align = 'l',
      format.args = list(big.mark = ',')) %>% 
  kable_styling()
```

As we can see from the table above, the proportion of population which is within the 100-mile zone seems to be in line with the ACLU, whereas the population within this zone seems to be higher than their number, suggesting there may be double countin of some population values. 

## Mapping the 100-Mile Zone

<center>

```{r, warning = F}
us_border_100 <- city_boundary %>% filter(dist_to_border <= 160)

top_10_per_state <- us_border_100 %>% group_by(state_name) %>% slice_max(population, n = 1)

mile_100_plot  <- ggplot() + 
  geom_sf(data = conus) +
  geom_sf(data = city_boundary, aes(col = dist_to_border), size = .05) +
  geom_sf(data = top_us_border)+
  geom_sf(data = top_10_per_state) +
  scale_color_gradient(low = 'orange', high = 'darkred') +
  gghighlight(dist_to_border <= 160, unhighlighted_colour = alpha("black", 0.4)) + 
  geom_label_repel(data = top_10_per_state ,
                   aes(label = city, geometry = geometry),
                   stat = 'sf_coordinates',
                   size = 2.75) +
  theme_map() +
  labs(title = 'Map of US Cities and 100-Mile Zone',
       subtitle = 'Most populous city in each state within the 100-mile zone labeled',
       col = 'Distance to \nBorder (km)')+ 
  theme(legend.key.size = unit(0.5, "cm"))
  

mile_100_plot
```
</center>

# Code Appendix

```{r, ref.label=knitr::all_labels(),echo=T,eval=FALSE}

```











